---
title: "Inference for repeated cross-validated AUC/concordance" 
execute: 
  echo: false
  cache: true
  warning: false
  message: false
  eval: true
format: 
  revealjs:
    reference-location: document
    theme: dark
    incremental: true
    progress: true
author: "Lily Koffman" 
date: "11 October 2024"
institute: "WIT"
---


```{r}
#| include: false
#| eval: true
library(gt)
library(tidyverse)
`%notin%` = Negate(`%in%`)
library(paletteer)

# custom color palette
# paletteer::paletteer_d("colorBlindness::paletteMartin")
coladept = "#DB6D00FF"
colsdt = "#920000FF"
coloak = "#FF6DB6FF"
colacti = "#490092FF"
colvso = "#006DDBFF"
colvsr = "#6DB6FFFF"
colscr = "#004949FF"
colscs= "#009292FF"


```


## Background

+ k-fold cross validation is standard procedure for estimating performance of model (e.g. AUC, concordance) 
+ One run of k-fold cross validation can create noisy estimate of model performance due to variation in splits of the data, especially in small datsets 
+ Repeated k-fold cross validation (simpy repeating the cross validation many times) leads to more stable estimates of model performance^[https://jcheminf.biomedcentral.com/articles/10.1186/1758-2946-6-10] 
+ Drawback: computation time 


## Currently

+ We have used 100-times repeated 10-fold cross validated concordance (rcvC) in previous papers
+ [Evaluating the prediction performance of objective physical activity measures for incident Parkinson’s disease in the UK Biobank](https://link.springer.com/article/10.1007/s00415-023-11939-0)
+ **Question**: how to do inference on concordance - i.e. how to test if one model is better than another?
+ Can we get a 95\% CI for the rcvC? 

## Intuition

+ rcvC is a random variable, with a distribution of values
+ We don't know the distribution of rcvC, but we can estimate it by bootstrapping
+ Resample subjects with replacement to create bootstrapped datasets
+ On each bootstrapped dataset, calculate rcvC 
+ Repeat *B* times 
+ Get 95\% CI for rcvC from bootstrapped values 

## In practice

+ We are not aware of this being done 
+ We are not sure that this leads to a valid 95\% CI for the concordance 

## Plan 
+ Simulate population of 1,000,000 data with known AUC and prevalence, since unclear on how to simulate data with known concordance  
+ Take sample from population 
+ Calculate 100x repeated 10-fold cross-validated AUC 
+ Create *B* bootstrapped samples from population
+ Calculate 100x repeated 10-fold cross-validated AUC on each bootstrapped sample
+ Use distribution of bootstrapped AUCs to get 95\% CI for population AUC 

## Simulation details 
+ Simulate from distribution with known AUC 
+ [Transforming the Area under the Normal Curve (AUC) into Cohen’s d, Pearson’s r pb , Odds-Ratio, and Natural Log Odds-Ratio: Two Conversion Tables](https://www.researchgate.net/publication/322205695_Transforming_the_Area_under_the_Normal_Curve_AUC_into_Cohen's_d_Pearson's_r_pb_Odds-Ratio_and_Natural_Log_Odds-Ratio_Two_Conversion_Tables) provides formula for converting AUC into Cohen's *d*
+ Cohen's *d*: difference between two group means in terms of standard deviations 
+ Simulate $n * (1-\text{prevalence})$ observations for non-events from $N(0, 1)$ and $n * \text{prevalence}$ observations for events from $N(d, 1)$


## Formulas

$$ t= \sqrt{ln\frac{1}{(1-AUC)^2}} $$
$$ z = t - \frac{2.515517 + 0.802853t + 0.0103328t^2}{1 + 1.432788t + 0.189269t^2 + 0.001308t^3} $$

$$ d = z \sqrt{2} $$

---

## Code 

```{r}
#| echo: true
#| eval: false 
sim_known_auc = function(auc_target, n, prevalence) {
  t <- sqrt(log(1 / (1 - auc_target) ** 2))
  z <- t - ((2.515517 + 0.802853 * t + 0.0103328 * t ** 2) /
              (1 + 1.432788 * t + 0.189269 * t ** 2 + 0.001308 * t ** 3)
  )
  d <- z * sqrt(2)
  x_events = rnorm(n * prevalence, mean = d)
  x_nonevents = rnorm(n * (1 - prevalence), mean = 0)
  y_events = rep(1, n * prevalence)
  y_nonevents = rep(0, n * (1 - prevalence))
  return(tibble(
    predictor = c(x_events, x_nonevents),
    outcome = c(y_events, y_nonevents)
  ))
  
}
```


## Parameters

+ Sample size: $\{1000, 5000, 10000, 100000\}$
+ Population AUC: $\{0.6, 0.7, 0.8, 0.9\}$
+ Population prevalence: $\{0.05, 0.1, 0.2, 0.3, 0.4, 0.5\}$


